---
title: "Statistics"
author: "Federspiel Sven"
date: '2021-11-19'
output:
  pdf_document: default
  html_document: default
---
  
```{r setup, include = FALSE}
install.packages("quantmod")
install.packages("goft")
install.packages("fitdistrplus")
install.packages("fitdistrplus")
library(quantmod)
library(tidyverse)
library(goft)
library(fitdistrplus)
```
  



## Import data

Command to retrieve data from yahoo finance, returns a time series object.

```{r}
getSymbols("AAPL", from = "2017-8-27", to = "2017-11-27", auto.assign = TRUE)

close <- AAPL$AAPL.Close
open <- AAPL$AAPL.Open
price <- AAPL$AAPL.Close
```


## Normalizing

```{r}
for (i in 0:length(close)){
  price[i] <- close[i]/open[i]
}

```


## extracting a numeric vector.

```{r}
x <- data.frame(coredata(price))
```

## Statistics.

Looking at empirical data and fitting a lnorm density curve to it.
It uses the MLE.

```{r}
plotdist(x$AAPL.Close, histo=TRUE, demp = TRUE)
fun1 <- fitdist(x$AAPL.Close, "lnorm")
plot(fun1)
```

Note : I expect these kind of figures to appear in the presentation, aswell
as a table with the test results from the next command

## testing log normality

(let's say we want p>0.1), if p is smaller we reject the
Lognormal hyp.

```{r}
lnorm_test(x$AAPL.Close)
```

## things to do and discuss

Selecting the stocks (as we are considering growth stocks I think taking the
top stocks from the nasdaq, maybe we can also check how indexes behave), also
we need to choose a number of stocks to work on 

Here we deal with day to day data, it would be nice to do the same with weekly
and or monthly data (either find a way to directly get the data from quantmod
or play with the indices from the day to day data.)

in the fitdistrplus package there's a goodness of fit command that would be
interesting to use, we shall have a look into it

Still some blabla to do for the introduction and structure of the presentation
we will discuss it in person.

priority is to check the fitdistrplus package and the last paper from section
4 till the end so you know what to expect. Also Im not even sure if the
way i normalize is correct, it seems natural but if you have anything against
or to add please do.
